### 2022-12-15 Manuscript.qmd
Update Manuscript.qmd
revision

---
title: "When Precision Therapy Creates Inprecise Choices"
subtitle: "Antifungal prophyalxis in the era or targeted therapy for acute myelogeneous leukemia"
format: html
editor: visual
bibliography: references.bib
csl: clinical-infectious-diseases.csl
---

Russell E. Lewis, Pharm.D.^1^

Marta Stanzani, M.D., Ph.D.^2^

^1^Associate Professor of Medicine, Department of Molecular Medicine, University of Padua, Padua, Italy:ORCID ID 000-0002-2002-4339

^2^Director of Hematopoietic Stem Cell Transplantation and Cellular Therapy,   Hematology Unit, Ca' Foncello Hospital,  AULSS 2- Marca Trevigana, Treviso, Italy:ORCID ID:000-0003-1569-3447

# Abstract

(150-250)

\* corresponding author: russ.e.lewis\@gmail.com

# Keywords:

# Competing interests:

Authors are required to disclose financial or non-financial interests that are directly or indirectly related to the work submitted for publication. Please refer to "Competing Interests and Funding" below for more information on how to complete this section.

# Introduction

Acute myeloid leukemia (AML) is a hematologic malignancy characterized by the proliferation of myeloid blasts or progranulocytes that fail to undergo normal differentiation [@newell_cook21]. It is the most common acute leukemia is older adults and historically associated with 5-year survival rates less than 10-15% in patients over 60 years of age [@kantarjian_etal21; @kantarjian_etal21b]. Traditional remission-induction chemotherapy (RIC) approaches for AML used for over 40 years such as 7 days of cytarabine + 3 days of daunarubin (7+3 regimen) cause severe and prolonged neutropenia, which places patients at high risk for invasive fungal diseases (IFDs) [@taplitz_etal18]. Beyond their direct mortality, fungal infections can jeopardize the timing and intensity of subsequent chemotherapy treatment cycles [@girmenia2014a; @even2011] or stem cell transplantation [@cordonnier_etal95]increasing the patient's risk for leukemia relapse. As a result, primary antifungal prophylaxis, first with fluconazole, and now more commonly with a mold-active triazole posaconazole, has become a standard of care in many centres [@maertens_etal18; @cornely2007J; @halpern2015].

Progress in the understanding of the biology and therapy of AML has evolved at a rapid pace over the last two decades. Advances in cytogenetic analysis and next-generation sequencing have better defined the molecular drivers of leukemogenesis and disease prognosis sparking an explosion in the development of "targeted therapies" [@kantarjian_etal21b]. Since 2017, nine new targeted therapies have been approved by the U.S. Food and Drug Administration of the treatment of AML with many additional agents in clinical trials. These therapies are now being incorporated in virtually every treatment phase of AML, and are increasingly are used in place of more toxic standard chemotherapy regimens in older adults who in the past were only candidates for palliative care or hospice care [@kantarjian_etal21b; @kantarjian_etal22].

Most of the recently approved targeted therapies are small molecule kinase inhibitors (SMKIs) that are metabolized through the cytochrome P450 3A4/5 oxidative enzymes in the liver [@bruggemann_etal22]. As such, their clearance in altered by varying degree by triazole antifungals potentially resulting in higher drug exposures to the targeted chemotherapy agent and potentially serious toxicities. This creates a dilemma for clinicians of whether to continue to use posaconazole prophylaxis, which has a proven survival benefit in AML patients with prolonged neutropenia, but would potentially put the patient at greater risk for treatment-limiting toxicities that may increase the risk of leukemia relapse.

Specifically, four scenarios are often considered: (i) stopping triazole prophylaxis to avoid interactions with the SMKI; (ii) switching to an alternative antifungal prophylaxis without CYP3A4/5 interactions; (iii) empirically reducing the dose of the SMKI to account for the effect of triazole CYP3A4/5 inhibition; and (iv) continuing both the SMKI and triazole antifungal dose and monitoring the patient carefully for toxicity. Each of these choices create risks to patient and must be considered individually depending on the treatment scenario and resources for patient monitoring, fungal diagnostic testing and therapeutic drug monitoring.

In this review, we address these four questions for the three most common class of prescribed SMKIs-the anti-apoptotic B/cell lymphoma-2 (Bcl-2) protein inhibitors venetoclax, fms like tyrosine kinase 3 (FLT-3) inhibitors (midostaurin, gilteritinib, quizartinib) and isocitrate dehydrogenase 1 and 2 (IDH1/2) inhibitors (ivosidenib, enasidenib). We will specifically focus on recent data examining what is known about the potential risks for IFD with SMKIs used in AML treatment that define the need for prophylaxis, what are the advantages and disadvantages of using non-triazole based prophylaxis, what is known about the safety of empiric dose adjustments for SMKIs, and what the are the risks giving triazole with full-dose SMKI therapy.

# Can primary antifungal prophylaxis during remission-induction chemotherapy with triazoles be stopped?

The use of antifungal prophylaxis is typically justified if the infection is (1) Serious enough such that even when diagnosed early or treated preemptively the infection incurs significant morbidity and mortality; (2) considered common enough to justify universal prophylaxis of the at risk population (i.e. \> 10% incidence), (3) the infection is difficult to diagnose. The first criteria is generally accepted to be true for IFDs in AML patients. The second criteria of how common the infection is , will vary from center to center and individual patient characteristics. Although recent surveys have suggested the incidence of IFDs in patients undergoing treatment for AML typically range between 10 to 25%, there is significant variation variation from one center to the next [@pagano_etal16] Nucci and Anaissie [@nucci_anaissie14] summarized the key risk factors for developing IFD in patients undergoing treatment for AML into 4 categories:

-   Host fitness for standard chemotherapy (i.e., fit, unfit, or frail);
-   Leukemia resistance (high versus low probability of achieving complete remission \[CR\]) and current malignancy status
-   Anticipated treatment-related toxicity such as neutropenia, mucositis, and steroid-induced immunosuppression
-   Patient exposure to opportunistic fungi.

While these categories are still relevant, use of SMKIs and hypomethylating agents instead of intensive cytotoxic chemotherapy, especially in older patients, may alter risk for invasive fungal disease [@aldoss_etal19]. Therefore, a central question is whether the use of targeted therapies is whether their use augments or reduces the risk of IFD in patients undergoing treatment for AML.

This question was the subject of recent reviews [@reinwald2018], and an expert guideline written by experts on behalf of the European Hematology Association (EHA) in cooperation with the Cochrane Hematology Group [@stemler_etal22]. For the resulting guidelines, studies including 19 relevant novel targeted agents currently approved or in later stages of clinical trials were reviewed, summarized and depending on available evidence, prepared as consensus-based recommendations despite the lack of high-quality studies assessing the use of antifungal agents in this setting [@stemler_etal22; @stemler_cornely22]. A summary of the recommendations for key approved agents is provided in [@tbl-drugs], the continued use of posaconazole prophylaxis was still recommended if patients are receiving targeted therapies as part of standard RIC. However, in consolidation chemotherapy, the risk of IFD is generally lower especially for invasive aspergillosis unless the patient has persistent neutropenia or prior fungal infections [@pagano_etal16], therefore continued use of posaconazole prophylaxis in patients received targeted therapies as maintenance or consolidation chemotherapy is conditional on the patient-specific risks. In this respect, we have found individualized risk stratification tools, such as institutionally-validated risk scores for aspergillosis [@stanzani_etal19; @stanzani_lewis18]to be helpful for decisions of whether continued prophylaxis is necessary.

The third key criteria supporting the use of prophylaxis is diagnostic difficulty. The availability of improved diagnostic tools (serum galactomannan, PCR) and high-resolution CT scanning has improved the early diagnosis of these infections allowing for possible or a preemeptive treatment approaches- i.e. initiation of antifungal therapy after detection of early fungal biomarkers and suggestive chest CT findings. Preemptive management approaches for IFD have been shown to be as effective as empirical treatment in high-risk patients with persistent fever [@morrissey2013]. An EORTC-lead randomized trial demonstrated that preemptive antifungal strategy was safe for high-risk neutropenic patients given fluconazole as prophylaxis, halving the number of patients receiving mould-active antifungals without excess mortality or IFDs [@Maertens2022]. However, primary antifungal prophylaxis with posaconazole has not been directly compared with empiric or preemptive therapy in a randomized clinical trial during RIC for AML.

In allogeneic HSCT recipients, a preemptive-like approach (fluconazole prophylaxis plus galactomannan monitoring) was proven to be as safe and effective as voriconazole prophylaxis both in terms of 6-month fungal-free survival (FFS) and relapse-free and overall survival [@Wingard2010a]. However, in patients transplanted for AML, patients who received voriconazole prophylaxis versus the preemptive approach experienced significantly fewer cases of IFD (8.5% vs. 21%, P=0.04) and had improved FFS (78% vs. 61%, P=0.04) and a trend towards improved overall survival (81% vs. 72%, P=0.32). Therefore, the equivalence of preemptive versus prophylactic approach for the highest groups remains unproven [@Coussement2021]. However, the safety of a preemptive approach in lieu of continued prophylaxis seems feasible in patients on SMKIs during consolidation phase, provided there is rapid availability to CT scanning and mycological testing, including when patients are in the outpatient setting where the bulk of these treatments are now being administered.

Collectively, the evidence does not seem to currently support abandoning primary antifungal prophylaxis in AML patients receiving SMKIs during RIC when the risk of IFD is highest. However, primary antifungal prophylaxis mold-active triazoles may be unnecessary for patients receiving lower- intensity therapy with a hypomethylating agent (decitabine or azacitidine) or low-dose cytarabine alone combined with drugs such as venetoclax unless extenuating risk factors are present, such as prolonged neutropenia [@nucci_anaissie14] In these patients a preemptive or empiric management approach may be appropriate in lieu of continued prophylaxis to reduce the risk of toxicity and drug interactions with the SMKIs.

# Should the patient be switched to alternative antifungal prophylaxis without CYP3A4/5 interactions?

Non-triazole based agents, namely echinocandins or liposomal amphotericin B, are often considered as alternatives to triazoles in patients who require SMKIs or are enrolled in clinical trials with new targeted therapies that preclude use triazoles because of QT-prolongation risk [@bruggemann_etal22]. Currently, only micafungin is approved for prophylaxis of Candida infection in patients undergoing allogeneic hematopoietic stem cell transplantation or patients who are expected to have neutropenia (absolute neutrophil count \< 500 cells / µl) for 10 or more days. Current guidelines for the treatment of invasive aspergillosis do not recommend echinocandin monotherapy as a frontline treatment for invasive aspergillosis due to limited evidence of efficacy in patients with probable or proven IA [@ullmann2012a; @viscoli2009; @herbrecht2010a]. A recent open-label randomized trial comparing caspofungin to fluconazole prophylaxis in children and adolescents undergoing treatment for AML found significantly lower cumulative incidence of proven or probable aspergillosis in patient who received caspofungin versus fluconazole (0.5% vs. 3%, P=0.04) [@fisher_etal19a], although the rates of IA in the fluconazole arm were lower than reported in previous adult studies that compared posaconazole suspension to fluconazole/itraconazole (1% vs. 7%, P=0.04) [@cornely2007J]. Some [@gomes2014] but not all [@Epstein2018a];El Cheikh et al., 2019, #101153} single-center observational studies examining echinocandin prophylaxis during RIC for AML have reported higher rates of breakthrough IFD in patients receiving echinocandin prophylaxis versus triazoles. Therefore, it is still unclear whether echinocandin prophylaxis, despite the improved safety profile when administered with SMKIs, is as effective as posaconazole prophylaxis for preventing invasive mold disease.

Echinocandin prophylaxis may be associated with higher risk for breakthrough fungal infections caused by rare yeast {Stanzani et al., 2019, #287617;Gomes et al., 2013, #95041; El Zein et al., 2020, #229138} and intrinsically-resistant mold such as *Fusarium* spp. and Mucorales compared to mould-active triazoles {Gomes et al., 2013, #95041}. In centers or regions where these infections are more common, careful consideration should be given to early diagnosis of these breakthrough infections given their propensity for fulminant infections and rapid dissemination.

Intravenous liposomal amphotericin B (L-AMB) is occasionally used off-label as a prophylaxis agent, although the optimal dosing approach and efficacy is not well established. This approach was often considered adult patients with acute lymphoblastic leukemic (ALL), who frequently receive chemotherapy regimens containing vinca alkaloids, which are metabolized through CYP3A4 {Stafylidis et al., 2022, #235425}. Cornely et al. performed a double-blind multicenter Phase 3 study comparing twice-weekly prophylactic liposomal amphotericin B (L-AMB) at 5 mg/kg intravenously or placebo twice during remission-induction treatment {Cornely et al., 2017, #34343}. Rates of proven and probable IFD were similar between L-AMB and placebo (7.9% vs. 11.7) with similar mortality rates. Hypokalaemia (35.0% vs. 17.8%) and increased creatinine (\> 1.5x baseline, 3.0% vs. 0%) were significantly more frequent in patients who received L-AMB prophylaxis. It is unclear if the limited efficacy was a reflection of the infrequent dosing regimen- therefore, if L-AMB is considered it may be more judicious to administer a daily 3 mg/kg dose for a define period (i.e. \< 21 days) if in unique cases L-AMB prophylaxis is used in patients who cannot tolerate triazoles or echinocandins.

Both echinocandins and liposomal amphotericin B must be administered intravenously, making them more problematic to use in the outpatient setting in patients taking SMKIs as part of a consolidation- maintenance regimen. An oral echinocandin-like triterpenoid recently approved for the treatment of vuvlvovaginal candidiasis, ibrexafungerp, is currently under investigations

If a triazole must be given, is isavuconazole the preferred?

What will be the potential role of new antifungals?

Table

Should the dose of the SMKI by empirically reduced to account for the effect of triazole CYP3A4/5 inhibition?

Venetoclax

FLT3 inhibitors

IDH1/2 inhibitors

What is the role of therapeutic drug monitoring?

## Conclusions

+-------------------------------------------------------+-----------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Targeted therapy                                      | Molecular target                                                                        | Mechanisms of immunosuppression risk                                                  | Impact of invasive fungal disease                                                                                                                                         | Triazole prophylaxis recommendation                                                                                                                                                                                                                   |
+=======================================================+=========================================================================================+=======================================================================================+===========================================================================================================================================================================+=======================================================================================================================================================================================================================================================+
| **Hypomethylating agents** (azacitabione, decitabine) | Inhibition of DNA methyltransferase that hypermethylase tumor suppressor gene promoters | Mild neutropenia and gastrointestinal toxicity                                        | Reported incidence in clinical studies 2.6-14.4%;                                                                                                                         | Conditional depending if neutropenia present at onset of use, or when used in combination with venetoclax. No dosage adjustment required when given with triazole.                                                                                    |
|                                                       |                                                                                         |                                                                                       |                                                                                                                                                                           |                                                                                                                                                                                                                                                       |
|                                                       |                                                                                         |                                                                                       | Low risk if used alone, but risk is higher when used in combination with venetoclax or in patients with prior neutropenia.                                                |                                                                                                                                                                                                                                                       |
+-------------------------------------------------------+-----------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Gemtuzumab-ozogamacin**                             | Humanized CD33-directed monoclonal antibody drug conjugate                              | Neutropenia, typically in combination with cytotoxic remission-induction chemotherapy | Limited evidence of increased risk for IFD above conventional remission-induction chemotherapy                                                                            | Triazole prophylaxis recommended when drug is used in combination for remission-induction chemotherapy. No dosage adjustment required when given with triazole.                                                                                       |
+-------------------------------------------------------+-----------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **BCL-2 inhibitor**                                   | Selective inhibitor of anti-apoptosis protein BCL2                                      | Neutropenia                                                                           | Limited evidence but overall risk is low, except in patients with prior allogeneic stem cell transplant, prolonged neutropenia, or when used for resistant/refractory AML | Conditional, but should be considered for patients with prolonged neutropenia. Dosage adjustment required when given with triazole                                                                                                                    |
|                                                       |                                                                                         |                                                                                       |                                                                                                                                                                           |                                                                                                                                                                                                                                                       |
| (venetoclax)                                          |                                                                                         |                                                                                       |                                                                                                                                                                           |                                                                                                                                                                                                                                                       |
+-------------------------------------------------------+-----------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **FLT3 inhibitors** (midostaurin, gilteritinib)       | Inhibition of fms like tyrosine kinase 3                                                | Neutropenia                                                                           | Limited evidence, no studies to date suggesting increased risk                                                                                                            | Triazole prophylaxis recommended when given as part of remission-induction chemotherapy. Individualized decisions when given as part of maintenance chemotherapy. Patients must be monitored closely for toxicities due to expected drug interactions |
+-------------------------------------------------------+-----------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| I**DH1/2 inhibitors**                                 | Isocitrate dihydrogenase i/ 2 enzyme inhibitors                                         | Toxicity                                                                              | Limited evidence, no studies to date suggesting increased risk                                                                                                            | Conditional, but should be considered for patients with prolonged neutropenia.                                                                                                                                                                        |
|                                                       |                                                                                         |                                                                                       |                                                                                                                                                                           |                                                                                                                                                                                                                                                       |
| (ivosidinib, enasidenib)                              |                                                                                         |                                                                                       |                                                                                                                                                                           |                                                                                                                                                                                                                                                       |
+-------------------------------------------------------+-----------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

: EHA Guidelines on Antifungal Prophylaxis in AML in Adult Patients Treatment with Novel Targeted Therapies.n {#tbl-drugs}